---
title: "Statistics on Illiteracy and Rape crime for US State"
author: "kjarnutowska"
date: "19 października 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this exercise I am working with two datasets. Data sets related to the 50 states of the United States of America. Both are from `datasets` package. 
First `state.x77` is giving some basics statistics like **Population** (population estimate as of July 1, 1975), **Income** (per capita income, 1974), **Illiteracy** (1970, percent of population), **Life Exp** (life expectancy in years 1969-71), **Murder** (murder and non-negligent manslaughter rate per 100,000 population, 1976), **HS Grad** (percent high-school graduates, 1970), **Frost** (mean number of days with minimum temperature below freezing (1931–1960) in capital or large city) and **Area** (in square miles).
Second is `USArrests`. This data set contains statistics (**Murder**, **Assault**, **UrbanPop**, **Rape**), in arrests per 100,000 in each of the 50 US states in 1973. 

## 3 biggest states from choosen ones (with area limit)

```{r warning=FALSE, message=FALSE}
library(datasets)
data(state) 

library(dplyr)
```
Working on `state.x77` I am changing data set class from matrix to data frame. 

```{r results='hide'}
state.x77 <- as.data.frame(state.x77)
state.x77$State <- rownames(state.x77) 
summary(state.x77) # summary to chcek if there are NaN's
```
Now I want to filter states with area between 50000 and 200000. Next is to select names of tree the biggest ones.
```{r}
stanyZPowierzchniaMiedzi50ka200k = state.x77 %>% 
    filter(Area >=50000 & Area <=200000) %>% 
    arrange(desc(Area)) 
    stanyZPowierzchniaMiedzi50ka200k %>% 
    select(State) %>% 
    head(3) 

```

## Rape Factor
Based on previous exercise I want to get **5 State**'s names with the lowest Rape ratio (Rape is provided by `USArrests` data set).
```{r}
data("USArrests") 

USArrests$State <- rownames(USArrests) 
```
```{r}
stanyWgWspPrzestepczpsciSeksualnejRosnaco <- USArrests %>% 
  filter(State %in% stanyZPowierzchniaMiedzi50ka200k$State) %>% # filter the states based on previouse exercise (area between 50000 and 200000)
  arrange(Rape)     # data set is Rape order, increasing
  
stanyWgWspPrzestepczpsciSeksualnejRosnaco %>% 
  select(State) %>% # Selecting State variable
  head(5)           # 5 states with the lowest Rape

```

Working on **stanyZPowierzchniaMiedzi50ka200k** mutate new boolen variable with information if the State Income exceed 4600 or not.

```{r}
stanyZPow50ka200k_filtr<- stanyZPowierzchniaMiedzi50ka200k %>% 
  mutate(has_income = as.character(Income>4600)) # new variable with income grater than 4600 (boolen type, values: TRUE / FALSE)
```
## Including Graphics
```{r warning=FALSE, echo = FALSE}
library(ggplot2)
```
```{r echo = FALSE}
 ggplot(stanyZPow50ka200k_filtr,aes(x = reorder(State,Illiteracy))) + # zmienna State zostaje uszeregowana malejąco funkcją reorder()
  geom_bar(aes(weight = Illiteracy, fill = has_income)) + # fill = has income daje nam dwa zbiory (kolory), ustalenie weight = Illiteracy pozwala na zliczanie wsp. analfabetyzmu
  labs(title = "US States illiteracy vs Income group", x = "", y ="illiteracy (1970, percent of population)") + #tytuł wygresu i osiek 
  scale_fill_manual("has_income", values =  c("#000066", "#377EB8"), labels = c("0", "1"))+ #deklaracja jak chcemy żeby syświetlony został nasz podział na 2 zbiory (kolorystyka i opis legendy)
  coord_flip() # zamieniamy oś OX z OY, np. dla czytelności
```

